Large-scale harmonic patterns clustering

### 2010-02-26 Paper/Paper.tex
iproved

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
Barrington et al. recently studied music texture using a set of ordered
MFCC vectors \cite{Barrington2009a}. 
Nearest neighbor for music also includes \cite{Cano2004} and ...

@@

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate. Practise showed us that they are reasonable, and the
size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments}).

@@

\section{Algorithm}\label{sec:algo}

Vector quantization \cite{Gersho1991}. Online learning.

\subsection{Codebook Learning}
We initialize the codebook by choosing $K$ random points from our dataset.

@@

\subsection{Setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $1e-2$ for $20$ iterations over the whole dataset.

@@

\subsection{Visualization}

>>>>

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.

Barrington et al. recently studied music texture using a set of ordered
MFCC vectors \cite{Barrington2009a}. This could be seen as complementary
to our work. Chromas capture the harmonic information and some of the
rythm while the MFCC are great timbral features.

Nearest neighbor for music also includes \cite{Cano2004} and ...

@@

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (e.g. in \cite{Barrington2009a} authors improve on the
song segmentation features). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments}).

@@

uspop2002 serves as testing set to measure how well a code book learned on
the Cowbell data set can represent new songs.

\section{Algorithm}\label{sec:algo}
In this section we present the creation of the harmonic patterns from
the Echo Nest data. Then, we explain the online vector quantization algorithm
used to learn a code book of those patterns.

\subsection{Pattern Creation}
Here we should explain key invariance, resizing, etc, and not in the
experiment section.

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online Kmeans.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained as Algorithm \ref{algo:vq}. note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

We initialize the codebook by choosing $K$ random points from our dataset.

@@

\subsection{Setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $0.01$ for $20$ iterations over the whole dataset.

@@

\subsection{Visualization}
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. We use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.
